GPU-based Adaptive Beamforming for Active Sonar Imaging

### 2012-05-08 ECUA2012/gpu_capon_ecua2012.tex
la2pdf commit - 2012:05:08	03:32:59	(Tue 8.May)

\maketitle

@@

As Figure \ref{benchmarks} illustrates, using a \gls{GPU} improves matters. If we look at the total runtimes for a subarray size of $L=16$ and $2N_\text{avg}+1=5$ pixels of temporal averaging, the \gls{GPU} implementation is able to process the test image in roughly 40\,ms. This translates to a processing throughput of 1.1\,Mpixels/s. For the same scenario the C implementation clocked in at roughly 3\,s and Matlab at 17\,s, which is 75 and 435 times slower than the \gls{GPU} implementation, respectively.

@@

@@

The \gls{MVDR} beamformer is an algorithm well suited for implementation on a \gls{GPU}. This is because the computations are independent on a pixel level, and partially also within each pixel. We were able to achieve speedup factors of 2-3 orders of magnitude by implementing our active sonar capable \gls{MVDR} implementation on a high-end \gls{GPU}. In sonar processing chains where the \gls{CPU} strain is considerable, simply being able to offloading some computationally intensive tasks to an idling \gls{GPU} might be just what is needed.

@@

\bibliographystyle{./template/ecua}
\bibliography{../../Library/library}

>>>>

\section*{Abstract}

Two limiting factors of modern phased array sonar imaging systems are the angular resolution and contrast. Adaptive beamformers target these limitatitons by weighting sensors dynamically to fit the impinging wavefield.

The Minimum Variance (MV) adaptive beamformer has recently been introduced in active sonar imaging. While displaying great potential for creating high quality images, it is computationally intensive and relies on robustification. These are important reasons why the MV beamformer is not in widespread use today.

We propose means for adapting the MV beamformer and its robustification techniques to fit a modern Graphics Processing Unit (GPU) architecture. These devices have in recent years
emerged as a flexible and highly potent parallel technology well suited for many DSP tasks, and are available off-the-shelf at reasonable prices.

We benchmarked optimized implementations of the MV beamformer in Matlab and C, running on a quad-core Intel Xeon E5620, and in CUDA running on a mid-range nVidia card. The beamformers operated on data from the 32x4 element Kongsberg Martime HISAS1030 sonar, using a conservative set of robustification parameters. We found that the GPU rendered more than 1Mpixels/s, representing a speedup of more than two orders of magnitude over C, and three orders of magnitude over Matlab. If the HISAS1030 was attached to a platform traveling at 1.8m/s, creating full-coverage sectorscan images in realtime would be possible with this performance. The CPU would then be free to take on the rest of the beamforming as well.

\newpage
\maketitle

@@

As Figure \ref{benchmarks} illustrates, using a \gls{GPU} improves matters. If we look at the total runtimes for a subarray size of $L=16$ and $2N_\text{avg}+1=5$ pixels of temporal averaging, the \gls{GPU} implementation is able to process the test image in roughly 40\,ms. This translates to a processing throughput of 1.4\,Mpixels/s. For the same scenario the C implementation clocked in at roughly 3\,s and Matlab at 17\,s, which is 75 and 435 times slower than the \gls{GPU} implementation, respectively.

@@

If the HISAS1030 was attached to a platform moving at 1.8m/s, for which the maximum range is 250\,m~\cite{Hansen2010}, the pulse repetition frequency could be set to 3\,Hz. Then the throughput required for creating critically sampled 32-beam sectorscan images in realtime would be roughly 1\,Mpixels/s. This is something our \gls{GPU} implementation could handle alone, leaving the \gls{CPU} free to take on other assignments.

@@

The \gls{MVDR} beamformer is an algorithm well suited for implementation on a \gls{GPU}. This is because the computations are independent on a pixel level, and partially also within each pixel. We were able to achieve speedup factors of 2-3 orders of magnitude by implementing our active sonar capable \gls{MVDR} implementation on a high-end \gls{GPU}. This performance is sufficient for computing critically sampled full-coverage sectorsscan images from the HISAS1030 sonar in realtime. Furthermore, offloading tasks to the \gls{GPU} might be desirable in itself since sonar processing chains are often \gls{CPU} intensive.

@@

\newpage
\bibliographystyle{./template/ecua}
\bibliography{../../Library/library}
